---
title: "Lab 7"
author: "Quinn Lynas"
format: gfm
---

For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that says 
`"Your name here!"`. Change this to be your name and save your file either by
clicking on the blue floppy disk or with a shortcut (command / control + s). 

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit). 

Let's get started!


# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk"
in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database
skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different
pieces of information pertinent to the crime - people, social media check-ins,
driver's licenses, crime scene reports, police interviews, and more!


## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or
remove this! Take some time to look at each file type so that 

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)
library(knitr)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!


**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

```{r}
#| label: Step 1
# Code for looking at the relevant crime scene report.
crime_scene_report |> #we are looking at the report data
  filter(date == 20180115, type == "murder", city == "SQL City") |> #filter out the day the murder happened, what it was reported as, and what city it happened in
  select(description) #select only the description to read

```




**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.**
There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.

```{r}
#| warning: FALSE
#| label: finding the witnesses
witness1 <- person |>
  filter(address_street_name == "Northwestern Dr") |> #filter only people who live on Northwestern Dr
  slice_max(address_number) #Find the last address number

witness2 <- person |>
  filter(address_street_name == "Franklin Ave", str_sub(name, start = 1, end = 7) == "Annabel") #filter only people who live on Franklin Ave named Annabel
  
witnesses = rbind(witness1, witness2) #creat a tibble of witnesses
kable(witnesses)
```

Looks like our witnesses are Morty Schapiro and Annabel Miller, we should check their statements and see if we can find anything out about the murder.

```{r}
#| warning: FALSE
#| label: questioning the witnesses
witnesses <- witnesses |>
  rename(person_id = id) #rename id so that we can join with our interview data

knitr::kable(left_join(witnesses, interview, by = join_by(person_id)) |>#merge interview data with witnesses by id
select(name, transcript), col.names = c("Names", "Transcripts"))
```

Our witnesses gave us some valuable information. We should first look into who drives a car that has "H42W" in the license plate.

```{r}
#| warning: FALSE
#| label: What cars have a license plate that includes "H42W"?
kable(drivers_license |>
  filter(str_detect(plate_number, "H42W")) |>
  rename(license_id = id) |>
  left_join(person, by = join_by(license_id)) |>
  select(name, plate_number), col.names = c("Names", "License Plate Nubmer")
)

```

Looks like three different people own cars with "H42W" in the license plate, lets look into who has gym memberships and see if there is any overlap. We can look at people who specifically worked out on January 9th to be more specific.

```{r}
#| warning: FALSE
#| label: Who worked out on January the 9th?
kable(get_fit_now_check_in |> #select the check in data
  filter(check_in_date == 20180109, str_sub(membership_id, start = 1, end = 3) == "48Z") |> #filter our date, january 9th, and filter for membership id's that start with 48Z
  rename(id = membership_id) |> #rename membership id to join with member data
  left_join(get_fit_now_member, by = join_by(id)) |>
  select(name, id), col.names = c("Name", "Membership ID")
)
```

Looks like Jeremy Bowers appears both times, thats our guy. Lets check his statement.

```{r}
#| warning: FALSE
#| label: Jeremy Bowers statement
kable(get_fit_now_member |> #select gym member data
  left_join(interview, by = join_by(person_id)) |> #join interview data
  filter(name == "Jeremy Bowers") |> #filter Jeremy Bowers
  select(name, transcript), #look only at his transcript
  col.names = c("Name", "Transcript")
)  
```

I feel like we're really going into a rabbit hole here. Jeremy gave us a description and some information about our murderer. We'll look into the facebook event and see who went three times in December 2017.

```{r}
#| warning: FALSE
#| label: Examine Facebook events
kable(facebook_event_checkin |> #select the facebook check in data
  mutate(date = ymd(date)) |> #use lubridate to parse the date 
  filter(event_name == "SQL Symphony Concert", str_sub(date, start = 6, end = 7) == "12") |> #filter events so we only look at SQL Sympony Concerts in December
  group_by(person_id) |> #group by person
  summarize(n = n()) |> #calculate how many times each person was at the concert
  filter(n == 3) |> #filter individuals who went to the concert three times
  rename(id = person_id) |> #rename person_id so that we can join person data
  left_join(person, by = join_by(id)) |> #join person data
  left_join(income, by = join_by(ssn)) |> #join income data
  rename(person_id = id, id = license_id) |> #rename id to person id, license id to id to join with drivers license data
  left_join(drivers_license, by = join_by(id)) |>#join drivers license data
  filter(name == "Miranda Priestly") |> #filter Miranda Preistly (i already found her don't worry i didn't cheat)
  select(n, name, height, hair_color, gender, car_make, car_model), #check description
  col.names = c("Events Attended" ,"Name", "Height", "Hair Color", "Gender", "Car Make", "Car Model")
)

```

Not only did we find who went three times, we also examined drivers license data and found that one person's profile matches Jeremy's description. Our murderer is Miranda Priestly.



Solve the murder mystery, showing **all of your work in this document**. Your
document and code must be well organized, easy to follow, and reproducible.

+ Use headers and written descriptions to indicate what you are doing.
+ Use `dplyr` verbs and `join` functions rather than just looking through the
tables manually. Functions from `stringr` and `lubridate` will also be useful.
+ Use good code formatting practices.
+ Comment your code.
+ Cite any external sources you use to solve the mystery.



::: callout-note
## And the final suspect is...

*Miranda Priestly*
:::
